I/O simulation model with Linux page cache, integration and evaluation in SimGrid framework

### 2020-12-07 paper_io_sim.tex
Replaced tabs by spaces

		In this section, we present our approach to model file read and write with 
		page cache, cache eviction and data flushing mechanisms implemented 
		in the Linux kernel. We also detail the design and implementation of two simulators:
		(1) a standalone Python prototype used to evaluate the model independently of any other 
		library, and (2) WRENCH-cache, an extension of WRENCH with our page cache model.
		Finally, we describe experimental scenarios to evaluate our simulations in various configurations.

@@

			We separate our simulation model in two components, the IO
			Controller and the Memory Manager, which together simulate 
			file read and write mechanisms in Linux (Fig.~\ref{fig:interaction}). 
			To read or write a file chunk, a simulated application sends a request 
			to the IOController. The IOController determines the amount of memory required by the operation,
			interacts with the Memory Manager accordingly, reads or write to cache through the Memory Manager, 
			and finally reads or writes on disk if necessary.
			The Memory Manager simulates flushing, eviction, and I/O to cache.

@@

			We leverage existing storage simulation models~\cite{lebre2015} to simulate the memory,
			characterized by its storage capacity, read and write bandwidths,
			and device latency.
			Leveraging existing storage models allows us 
			to take advantage of bandwidth sharing between concurrent I/Os to 
			simulate parallel memory accesses.

			\paragraph*{Memory Manager}
			The Memory Manager simulates  two parallel threads: the first and main one simulates 
			flushing, eviction, and cache I/Os synchronously; the second one is a background thread that  
			periodically searches for expired dirty data in page cache LRU lists 
			and flushes them to disk.

			\paragraph*{Page cache LRU lists}
			In the Linux kernel, the page cache LRU lists contain file pages. However, 
			due to the large number of file pages, managing lists of pages 
			induces substantial overheads.
			Thus, we introduce the concept of a data block as a unit to represent data 
			cached in memory. A data block is a subset of file pages stored in
            page cache that were accessed in the same I/O operation. 
			A data block has information about file name, block size, last access 
			time, a dirty flag that represents whether the data is clean (0) 
			or dirty (1), and the time it is written to cache as dirty data. 
			Blocks can have different sizes and a given file can have multiple 
			data blocks in page cache. In addition, a data block can be split into an 
			arbitrary number of smaller blocks.

>>>>

        In this section, we present our page cache simulation model for file I/O,
        flusing, and eviction. We detail the design and implementation of two simulators:
        (1) a standalone Python prototype to evaluate the simulation model independently of any other 
        tool, and (2) WRENCH-cache, an extension of WRENCH with our page cache model.
        We also describe experimental scenarios to evaluate our simulators in various configurations.

@@

            We separate our simulation model in two components, the IO
            Controller and the Memory Manager, which together simulate 
            file reads and writes in Linux (Fig.~\ref{fig:interaction}). 
            To read or write a file chunk, a simulated application sends a
            request to the IOController. The IOController interacts as needed with
            the Memory Manager to free memory through flushing or eviction,
            and to read or write to cache or to disk. The Memory Manager
            implements these operations, and simulates periodical flushing
            and eviction from cache.

@@

            We use existing storage simulation models~\cite{lebre2015} to simulate memory,
            characterized by its storage capacity, read and write bandwidths,
            and latency. Reusing existing storage models allows us 
            to simulate bandwidth sharing between concurrent memory accesses.

            \paragraph*{Memory Manager}
            The Memory Manager simulates two parallel threads: the main one
            implements flushing, eviction, and cache I/Os synchronously,
            whereas the second one, operating in the background,
            periodically searches for expired dirty data in page cache LRU
            lists and flushes them to disk.

            \paragraph*{Page cache LRU lists}
            In the Linux kernel, page cache LRU lists contain file pages. However, 
            due to the large number of file pages, managing lists of pages 
            induces substantial overheads.
            Thus, we introduce the concept of a data block as a unit to represent data 
            cached in memory. A data block is a subset of file pages stored in
            page cache that were accessed in the same I/O operation. 
            A data block has information about file name, block size, last access 
            time, a dirty flag that represents whether the data is clean (0) 
            or dirty (1), and the time it is written to cache as dirty data. 
            Blocks can have different sizes and a given file can have multiple 
            data blocks in page cache. In addition, a data block can be split into an 
            arbitrary number of smaller blocks.
